---
title: "Ch3_through_5_analyses"
output: word_document
date: "2024-06-17"
---

# Set Up 
## load packages
```{r}
library(dplyr)
library(tidyverse)
library(lme4)
library(lmerTest)
library(optimx) #to select an optimizer for glmer models
library(performance) #to check for collinearity between predictors
```
## load data 
These files contain both human and surprisal data. The Det/N data file contains data from both experiments (Chapters 3&4); experiment == 1 is Chapter 3 data, experiment == 2 is Chapter 4 data. 
```{r}
#det/n data
detn_data = read.csv("~/Desktop/data/processed/Ch5_detn_data_filtered.csv")
#adj/n cs between data (Chapter 3)
adjn_between = read.csv("~/Desktop/data/processed/Ch5_adjn_cs_between_data_filtered.csv")
#adj/n cs before data (Chapter 4)
adjn_before = read.csv("~/Desktop/data/processed/Ch5_adjn_cs_before_data_filtered.csv")
#frequency data of the adjectives and nouns in the adj/n stimuli
freq_data <- read.csv("~/Desktop/data/processed/adj_n_freq.csv")
```
## format data
```{r}
#det/n data
detn_data$experiment <-factor(detn_data$experiment)
detn_data$subject <-factor(detn_data$subject)
detn_data$starting_lang <-factor(detn_data$starting_lang)
detn_data$lang_switch <-factor(detn_data$lang_switch)
detn_data$cs <-factor(detn_data$cs, levels=c("CS", "noCS"))
detn_data$sensical <-factor(detn_data$sensical, levels= c("sense", "nonsense"))
detn_data$condition <- factor(detn_data$condition)
detn_data$response <- factor(detn_data$response)
detn_data$label <- factor(detn_data$label)

#adj/n cs between data
adjn_between$subject <-factor(adjn_between$subject)
adjn_between$starting_lang <-factor(adjn_between$starting_lang)
adjn_between$lang_switch <-factor(adjn_between$lang_switch)
adjn_between$cs <-factor(adjn_between$cs, levels=c("CS", "noCS"))
adjn_between$word_order <-factor(adjn_between$word_order, levels= c("AdjN", "NAdj"))
adjn_between$condition <- factor(adjn_between$condition)
adjn_between$response <- factor(adjn_between$response)
adjn_between$label <- factor(adjn_between$label)
adjn_between$mlwo <- factor(adjn_between$mlwo)
adjn_between$mlwo_condition <- factor(adjn_between$mlwo_condition)
adjn_between$cwh <- factor(adjn_between$cwh)
adjn_between$cwh_condition <- factor(adjn_between$cwh_condition)

#adj/n cs before data
adjn_before$subject <-factor(adjn_before$subject)
adjn_before$starting_lang <-factor(adjn_before$starting_lang)
adjn_before$lang_switch <-factor(adjn_before$lang_switch)
adjn_before$cs <-factor(adjn_before$cs, levels=c("CS", "noCS"))
adjn_before$word_order <-factor(adjn_before$word_order, levels= c("AdjN", "NAdj"))
adjn_before$condition <- factor(adjn_before$condition)
adjn_before$response <- factor(adjn_before$response)
adjn_before$label <- factor(adjn_before$label)
adjn_before$mlwo <- factor(adjn_before$mlwo)
adjn_before$mlwo_condition <- factor(adjn_before$mlwo_condition)
adjn_before$cwh <- factor(adjn_before$cwh)
adjn_before$cwh_condition <- factor(adjn_before$cwh_condition)
```
# Analyses
All analyses for each data set are run on multiple word positions, but the code for each analysis is only written once. Change the word number in "subset= word_no %in% c**#**" to the desired word position to run the analysis on the desired word position.
Word positions corresponding to the critical region and one word spillover region for all data sets: 
* Det/N: 5, 6
* Adj/N: 5, 6, 7
For models that do not converge, remove the three way interaction and instead include all 2-way interactions. 
* the 3-way interaction of *cs x sensical x starting_lang* (x=interaction) --> *cs x sensical + cs x starting_lang + sensical x starting_lang*
* the 3-way interaction of *cs x word_order x starting_lang* (x=interaction) --> *cs x word_order + cs x starting_lang + word_order x starting_lang*

## Analyses for Ch 3
### Det/N stimuli
```{r}
logistic <-glmer(response~cs*sensical*starting_lang+centered_span_prof+(1|subject)+(1|label), data=detn_data, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5) & experiment=="1")
summary(logistic)
```
### Adj/N CS Between Stimuli
```{r}
logistic <-glmer(response~cs*word_order*starting_lang+centered_span_prof+(1|subject)+(1|label), data=adjn_cs_between, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5))
summary(logisitic)
```
### CWH and MLF Analyses
CWH and MLF make the same predictions for all conditions, so the analysis is only run on one of the variables. 
```{r}
mlwo_logisitic <-glmer(response~mlwo+span_prof+(1|subject)+(1|label), data=adjn_cs_between, family=binomial, control=glmerControl(optimizer="bobyqa"), subset= word_number %in% c(5))
summary(mlwo_logisitic)
```

## Analyses for Ch 4
### Det/N Stimuli
```{r}
logistic <-glmer(response~cs*sensical*starting_lang+centered_span_prof+(1|subject)+(1|label), data=detn_data, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5) & experiment=="1")
summary(logistic)
```
### Adj/N CS Before Stimuli
```{r}
logistic <-glmer(response~cs*word_order*starting_lang+centered_span_prof+(1|subject)+(1|label), data=adjn_cs_before, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5))
summary(logisitic)
```
### CWH and MLF Analyses
#### MLF
```{r}
mlwo_logisitic <-glmer(response~mlwo+span_prof+(1|subject)+(1|label), data=adjn_cs_before, family=binomial, control=glmerControl(optimizer="bobyqa"), subset= word_number %in% c(5))
summary(mlwo_logisitic)
```
#### CWH
```{r}
cwh_logisitic <-glmer(response~cwh+span_prof+(1|subject)+(1|label), data=adjn_cs_before, family=binomial, control=glmerControl(optimizer="bobyqa"), subset= word_number %in% c(5))
summary(cwh_logisitic)
```

## If Necessary: Check which optimizers will allow models to converge
If needed, check which optimizer(s) will allow the model to converge. Create a logit model with all fixed and random effects (first chunk of code below), then run allFit command on the model (second chunk below).
```{r}
logistic_test <-glmer(response~cs*sensical*starting_lang+span_prof+(1|subject)+(1|label), data=detn_data, family=binomial, subset= word_number %in% c(7))
summary(logistic_test)
```
```{r}
diff_optims <- allFit(logistic_test, maxfun = 1e5)
```

## Analyses for Ch 5
### What is surprisal sensitive to?
#### Det/N stimuli
```{r}
regression <-lmer(word_surprisal~cs*sensical*starting_lang + (1|word), data=detn_data, subset= word_no %in% c(5))
summary(regression)
confint(regression)
```
#### Adj/N CS Between Stimuli
```{r}
regression <-lmer(word_surprisal~cs*word_order*starting_lang + (1|word), data=adjn_between, subset= word_no %in% c(5))
summary(regression)
confint(regression)
```
#### Adj/N CS Before Stimuli
```{r}
regression <-lmer(word_surprisal~cs*word_order*starting_lang + (1|word), data=adjn_before, subset= word_no %in% c(5))
summary(regression)
confint(regression)

#t-test to check if at word position 5 in non-codeswitched conditions, adjectives are more surprising in Spanish than in English
t.test(adjn_before$word_surprisal ~ adjn_before$starting_lang, subset = (adjn_before$word_no==5 & adjn_before$cs=="noCS" & adjn_before$word_order=="AdjN"), p.adjust="bonferroni")
```

### To what extent can surprisal predict human responses?
#### Det/N stimuli
```{r}
logistic <-glmer(response~word_surprisal+cs*sensical*starting_lang+(1|subject)+(1|label), data=detn_data, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5))
summary(logistic)
```
#### Adj/N CS Between Stimuli
```{r}
logistic <-glmer(response~word_surprisal+cs*word_order*starting_lang+(1|subject)+(1|label), data=adjn_before, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5))
summary(logistic)
```
#### Adj/N CS Before Stimuli
```{r}
logistic <-glmer(response~word_surprisal+cs*word_order*starting_lang+(1|subject)+(1|label), data=adjn_before, family=binomial, glmerControl(optimizer ='optimx', optCtrl=list(method='nlminb')), subset= word_number %in% c(5))
summary(logistic)
```

###	Comparing the MLF, CWH, and SCH: 
#### To what extent can the MLF or CWH predict human responses?
##### MLF
```{r}
logistic <-glmer(response~word_surprisal+mlwo+(1|subject)+(1|label), data=adjn_before, family=binomial, subset= word_number %in% c(5))
summary(logistic)
```
##### CWH
```{r}
logistic <-glmer(response~word_surprisal+cwh+(1|subject)+(1|label), data=adjn_before, family=binomial, subset= word_number %in% c(5))
summary(logistic)
```

####	To what extent can the SCH **and** MLF or CWH predict human responses?
##### SCH + MLF
```{r}
logistic <-glmer(response~word_surprisal+mlwo+(1|subject)+(1|label), data=adjn_before, family=binomial, subset= word_number %in% c(5))
summary(logistic)
```

##### SCH + CWH
```{r}
logistic <-glmer(response~word_surprisal+cwh+(1|subject)+(1|label), data=adjn_before, family=binomial, subset= word_number %in% c(5))
summary(logistic)
#check the collinearity of the SCH and CWH in the logisitic model
check_collinearity(logistic)
```

### Post-hoc Frequency Analysis
Compare the frequency of nouns and adjectives used in the adj/n stimuli.
```{r}
freq_test <-t.test(freq_per_mill~gram_cat, data=freq_data)
freq_test
```